---
title: "Permutation tests for adaptive designs"
subtitle: "Prototypes for API"
author: Florian Klinglmueller
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  \usepackage[utf8]{inputenc}
---


There are two ways to perform a test following an adaptive design:

1. Using a decision rule that compares the conditional error to
   the second stage p-value
2. Using an adjustde p-value which is then compared to the overall
   significance level



# Decision rules

Decision rules permit us to decouple computation of the conditional
error rate from the adapted second stage test. This simplifies
switching between different types of conditional error rate
computations.

We implement several ways to compute the conditonal error rate of the
pre-planned permutation test:

1. Non-randomized: this is the conditional error rate of the standard
   non-randomized permutation test. Depending on the discreteness of
   the permutation distribution this test will be strictly
   conservative. 
2. Randomized: this is the conditional error rate of the standard
   randomized permutation test. Which rejects for the same samples as
   the non-randomized test; on the border of the rejection region one
   rejects with probability $\gamma$ such that the test has size
   exactly $\alpha$.
3. Uniform: this is the conditional error rate of the non-randomized
   test plus the difference in size of the non-randomized and the
   significance level $\alpha$
   


```{r}
library(adaperm)


##' Permutation test p-value
##'
##' Type "non-randomized" will give the standard p-value, which is define by the proportion of sample permutations with larger or equal test statistic than the observed one. "midp" will give the mid-p value which is the proportion of sample permutations with larger test statistic plus one half the proportion of permutations with equal test statistic. "randomized" will return the proportion of permutations with larger test statistic plus a random number that is uniformly distributed between zero and the proportion of permutations with test statistic equal to the observed.
##' 
##' @title Permutation p-value
##' @param t 
##' @param dist 
##' @param type 
##' @return p-value
##' @author Florian Klinglmueller
t2p  <- function(t,dist,type=c('non-randomized','midp','randomized','davison_hinkley')){
    switch(type[1],
           'non-randomized'=mean(dist>=t),
           'midp'=mean(dist>t) + .5*mean(dist==t),
           'randomized'=mean(dist>t) + runif(1)*mean(dist==t),
           'davison_hinkley'=(sum(dist>t)+1)/(sum(dist!=t)+1))
}


##' Level alpha critical boundary of permutation test. Returns the smallest unique value of the permutation distribution such that the proportion of statistics larger is smaller than alpha. 
##' 
##' @title p to t
##' @param p p-value/alpha level
##' @param dist permutation distribution
##' @return critical value
##' @author Florian Klinglmueller
p2t <- function(p,dist)
    max(dist[rank(-dist,ties='min') >= p*length(dist)])
}




perm_test <- function (x1, x2, g1, g2, stat, B, x3 = NULL, g3 = NULL,
                       restricted,
                       type=c('non-randomized','randomized','midp','davison_hinkley')) 
{
    dist <- perm_dist(x1, x2, g1, g2, stat, B, x3, g3, restricted = restricted)
    t <- stat(c(x1, x2, x3), c(g1, g2, g3))
    t2p(t,dist,type[1])
}

permutation_cer <- function(x1,x2,
                            g1,nt2=floor(length(x2)/2),
                            test_statistic,
                            alpha,
                            permutations,
                            restricted,
                            cer_type=c("non-randomized","randomized","uniform")){
    n1 <- length(x1)
    n <- length(c(x1,x2))
    n2 <- n-n2
    g2 <- rep(0:1,each=nt2)
    pdist <- adaperm:::perm_dist(x1,x2,g1,g2,test_statistic,permutations,restricted=restricted)
    cdist <- adaperm:::cond_dist(x1,x2,g1,g2,test_statistic,permutations,restricted=restricted)
    M <- length(dist)
    talpha <- p2t(alpha,pdist)
    A  <- mean(cdist > talpha)
    trimmings <- switch(cer_type[1],
                        'non-randomized' = 0,
                        'randomized' = mean(cdist==talpha) * (alpha - mean(pdist > talpha)) / mean(pdist==talpha),
                        'uniform' = (alpha - mean(pdist > talpha)))
    A + trimmings
}
    
adpative_permdr <- function(x1,x2,xE,
                        g1,g2,gE,
                        test_statistic,
                        alpha,
                        permutations,
                        restricted,
                        atest_type=c("non-randomized","mid-p","davison_hinkley","CER"),
                        cer_type=c("non-randomized","randomized","uniform")){
    A <- permutation_cer(x1,x2,
                         g1,sum(g2>0),
                         test_statistic,
                         alpha,
                         permutations,
                         restricted,
                         cer_type)
    if(atest_type == 'CER'){
        return(A)
    } else {
        return(A >= perm_test(x2,xE,g2,gE,test_statistic,permutations,restricted=restricted,type=atest_type))
    }
}
.cer_types = c("non-randomized","randomized","uniform",)
.atest_types = c("non-randomized","mid-p","davison_hinkley","CER")

adaperm_DR <- function(x,g=NULL,n1,n,m1=n1,m=n,test_statistic,nt2=NULL,conf_level=.025,cer_type=.cer_types,atest_type=.atest_types,permutations=10000){
    if(is.null(g)){
        ## one-sample test
        restricted <- FALSE
        obs <- split_sample_os(x,n1,n)
    } else {
        restricted  <- TRUE
        if(length(g) == n1 && cer_type='CER'){
            ## if we only want to know the CER
            g <- c(g,rep(0,n-n1-nt2),rep(1,nt2))
        }        
        obs <- split_sapmle_ts(x[g<=0],y[g>0],n1,n,m1,m)
    }
    xs <- obs[[1]]
    gs <- obs[[2]]
    apaptive_permdr(xs[[1]],xs[[2]],xs[[3]],
                      gs[[1]],gs[[2]],gs[[3]],
                      test_statistic,
                      conf_level,
                      permutations,
                      restricted,
                      atest_type,
                      cer_type)
}
                      
                      
    
```

## Old vs. new API

```{r}

decision_rules <- data.frame(purpose =
                                 c('Compute CER',
                                   'general decision rule',
                                   'one-sample','two-sample',
                                   'user friendly'),
                             old_api =
                                 c(permutation_CER,
                                   adaptive_perm_test,
                                   list(adaptive_permtest_os,adaptive_DR_os),
                                   list(adaptive_pormtest_ts,adaptive_DR_ts),
                                   NA),
                             new_api =
                                 c(permutation_cer,
                                   adaptive_permtest,
                                   NA,
                                   NA,
                                   adaperm))
                             

```
